---
title: '**Advanced Macroeconometrics -- Assignment 1**'
author: 
  - "Siegfried Hammer (h12229325@wu.ac.at)"
  - "Max Heinze (h11742049@wu.ac.at)"
  - "Tim Koenders (h12215486@wu.ac.at)"
date: "April 19, 2023"
output: 
  pdf_document:
    includes:
      in_header: !expr file.path(rprojroot::find_rstudio_root_file(), "helper", "wrap_code.tex")
header-includes: \usepackage{tcolorbox}
papersize: a4
geometry: margin = 2cm
urlcolor: Mahogany
---

```{r, setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

\begin{tcolorbox}
\centering \itshape The executable code that was used in compiling the assignment is available on GitHub at \url{https://github.com/maxmheinze/macrometrics}.
\end{tcolorbox}

# Exercise 1

First, we read in the data set, removing the first column indicating the recommended transformation.

```{r, echo = TRUE, results = "hide"}
# Header ------------------------------------------------------------------

rm(list = ls())
gc()

pacman::p_load(
  tidyverse,
  urca
  )


# Read in Data ------------------------------------------------------------

fred <- read.csv("./assignment1/data/fred.csv")[-1,]
```

Next, we create the desired function `ts_explode()`. Along with the vector to be transformed, it asks for a specification of whether the data is ordered with the latest or earliest value first. It takes the earliest value first as default.

```{r, echo = TRUE}
# Create the function -----------------------------------------------------

ts_explode <- function(input_vector, start_with_latest = FALSE) {
  # The function is called ts_explode because a single vector explodes into an entire data frame. Boom!
  
  # Package dplyr required for lag() function
  require(dplyr)
  
  # Reverse input vector if user specifies it is sorted latest to earliest
  input_vector <- if (start_with_latest == FALSE) {
    input_vector
  } else {
    rev(input_vector)
  } 
  
  # Do the transformations, assign transformed vectors
  original <- input_vector
  log_transformed <- log(input_vector)
  mom_growth <- input_vector/dplyr::lag(input_vector, 1) - 1
  yoy_growth <- input_vector/dplyr::lag(input_vector, 12) - 1
  yoy_growth_lagged <- dplyr::lag(input_vector, 12)/dplyr::lag(input_vector, 24) - 1
  
  # Create a data frame to export, reverse ordering back to original in case start_with_latest = TRUE was specified
  export_df <- if (start_with_latest == FALSE) {
    data.frame(original, log_transformed, mom_growth, yoy_growth, yoy_growth_lagged)
  } else {
    data.frame(original = rev(original), 
               log_transformed = rev(log_transformed), 
               mom_growth = rev(mom_growth), 
               yoy_growth = rev(yoy_growth), 
               yoy_growth_lagged = rev(yoy_growth_lagged))
  }
  
  # Display warnings regarding ordering and units of growth rates
  warning("By default, ts_explode() assumes that values are ordered from earliest to latest. If your vector is ordered from latest to earliest, specify `start_with_latest = TRUE`!")
  warning("Growth rates are given in decimals, not in percent!")
  
  # Return the data frame
  return(export_df)
}
```

Using `ts_explode()`, we create the data frame `ind_prod` including all transformations of the `INDPRO` variable. We bind the data frame together with the date column, which we transform from character to date. All other changes in the resulting data frame are of cosmetic nature.

```{r, warning=FALSE}
# Prepare Industrial Production Data Frame --------------------------------

ind_prod <- fred$sasdate %>%
  cbind(ts_explode(fred$INDPRO)) %>%
  as_tibble() %>%
  mutate(date = lubridate::mdy(`.`)) %>%
  select(-`.`) %>%
  relocate(date, .before = original)
```

Next, we plot both the logged variable and the year-on-year growth rate.

```{r, echo = TRUE, out.width="50%", fig.align='center'}
# Create Log Plot ---------------------------------------------------------

ind_prod %>%
  ggplot() +
  geom_line(aes(x = date, y = log_transformed)) +
  labs(title = "U.S. Industrial Production (logged)",
       x = "Date",
       y = "Log of Industrial Production") +
  ylim(3,5) +
  theme_bw()


# Create Growth Plot ------------------------------------------------------

ind_prod %>%
  ggplot() +
  geom_line(aes(x = date, y = yoy_growth)) +
  labs(title = "U.S. Industrial Production (year-on-year growth)",
       x = "Date",
       y = "Year-on-Year Growth of Industrial Production") +
  theme_bw()
```

